Giant Low Surface Brightness Galaxies As Merger Remnants

### 2015-04-13 glsb.tex
a

\title{Giant Low Surface Brightness Galaxies as Merger Remnants}

\author{Peter Yoachim\altaffilmark{1}, Denise Shmitz\altaffilmark{1}, Sarah Loebman\altaffilmark{2}, SungWon Kwak\altaffilmark{1}, Victor P. Debattista\altaffilmark{3}
  }

\altaffiltext{1}{Department of Astronomy, University of Washington, Box 351580,
Seattle WA, 98195}
\altaffiltext{2}{XXX-U Mich}
\altaffiltext{3}{Jeremiah Horrocks Institute, University of Central Lancashire, Preston PR1 2HE, UK.  RCUK Fellow}

@@

Once the Vaccine routines are completed we make the following additional reductions using standard IRAF routines and custom IDL scripts.  The reduced twilight frames are used to generate a fiber-to-fiber flux correction (similar to a longslit illumination correction).  We use the b-spline procedure described in \citet{Kelson03} to generate background frames.  The corners of each frame contain enough empty fibers to measure the sky background in each exposure.  Because we bin fibers together, imperfect sky-subtraction represents a serious source of potential systematic errors.

@@

In addition to individual fibers, we have binned neighboring fibers together to boost the signal-to-noise ratio and run pPXF on these super-fibers.

@@

Figure~\ref{elliptical_relations} shows where the GLSBs fall on the major galaxy relations.  We use extinction corrected magnitudes and $k$-corrected colors\citep{Chilingarian10}.  Both of our GLSBs lie on the red-sequence in the color-magnitude diagram.

@@

\subsection{GLSB Disks}
The diameters of these disks are very large.  Stellar disks are typically embedded in much larger \HI\ disks.  The size of the GLSB stellar disks is similar to the largest \HI\ disks observed in nearby galaxies, in support of the \citet{Mapelli08} model where the GLSB disks are the faded remnants of a normal \HI\ disk that has experienced a merger induced star-burst.

For comparison:  \citet{Williams13} trace star formation out to 11 disk scale lengths (radius of 22 kpc) in NGC 2403.  Maybe also look at THINGS data for comparisons on how large galaxy disks get.

@@

In Figure~\ref{disk_comp}, we compare the sizes of the stellar disks of the GLSB galaxies with other observations.  For this plot, we take the maximum detected radius from the literature for the galaxies.  It is possible that deeper observations could increase the values of $R_{max}$.  While UGC 6614 close in size to the upper end of local \HI\ disks, Malin 1 and 2 are much larger than galaxies observed in the THINGS survey.  This seems to point against the GLSB formation mechanism of  \citet{Mapelli08}, as we do not observe \HI\ disks large enough to convert into GLSB systems without some additional source of angular momentum.

%/astro/store/shared-scratch1/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, Malin 1 from \citet{Moore06}, Malin 2 and UGC 6614 from \citet{Pickering97}.  Here, $R_{max}$ is the maximum radius with an \HI\ detection.  It is interesing to note that the stellar disks of GLSB galaxies are often larger than the \HI\ disks of nearby systems.  \label{disk_comp}}
\end{figure}

@@

%/astro/store/scratch/tmp/yoachim/Research/GLSB/THINGS/guess_projen.py
\begin{figure}
\epsscale{0.5}
\plotone{Plots/things_projenguess.eps}
\epsscale{1}
\caption{Comparison of GLSB \HI\ masses with \HI\ masses measured in the THINGS survey \citep{Walter2008}.  While the \HI\ masses in the GLSB systems are high, they are consistent with what would be available in a merger with a $\sim$110 km/s galaxy, assuming the merger did not trigger large amounts of star formation and feedback. \label{thingsMHI}}
\end{figure}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_Malin2.eps}{Plots/Malin2mu.eps}
\epsscale{.5}
\plotone{Plots/Malin2profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of Malin 2.  Upper Right:  Image made with the VIRUS-P datacube.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  }
\end{figure}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_UGC06614.eps}{Plots/UGC06614mu.eps}
\epsscale{0.5}
\plotone{Plots/UGC06614profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of UGC 06614.  Upper Right:  Image made with the VIRUS-P datacube.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  We only included fibers beyond the outer ring structure.  }
\end{figure}

>>>>

\title{Giant Low Surface Brightness Galaxies As Merger Remnants}

\author{Peter Yoachim\altaffilmark{1}, Denise Shmitz\altaffilmark{4}, Sarah Loebman\altaffilmark{2}, SungWon Kwak\altaffilmark{1}, Victor P. Debattista\altaffilmark{3}
  }

\altaffiltext{1}{Department of Astronomy, University of Washington, Box 351580,
Seattle WA, 98195}
\altaffiltext{2}{XXX-U Mich}
\altaffiltext{3}{Jeremiah Horrocks Institute, University of Central Lancashire, Preston PR1 2HE, UK.  RCUK Fellow}
\altaffiltext{4}{XXX-cal tech}

@@

Once the Vaccine routines are completed we make the following additional reductions using standard IRAF routines and custom IDL scripts.  The reduced twilight frames are used to generate a fiber-to-fiber flux correction (similar to a longslit illumination correction).  We use the b-spline procedure described in \citet{Kelson03} to generate background frames.  The corners contain enough empty fibers to measure the sky background in each exposure.  Because we bin fibers together, imperfect sky-subtraction represents a serious source of potential systematic errors.

@@

In addition to individual fibers, we have binned sets of three neighboring fibers together to boost the signal-to-noise ratio and run pPXF on these super-fibers. The fiber surface brightnesses are plotted in Figures~\ref{Malin2_fibers} and~\ref{ugc_fibers}, and the pPXF stellar velocity maps are shown in Figures~\ref{malin_kinem} and~\ref{ugc_kinem}.

@@

Figure~\ref{elliptical_relations} shows where the GLSBs fall on the major galaxy relations.  We use extinction corrected magnitudes and $k$-corrected colors \citep{Chilingarian10}.  Both of our GLSBs lie on the red-sequence in the color-magnitude diagram.

@@

\subsection{GLSB Disks}

The diameters of GLSB disks are very large, with Malin~1's stellar dist detected out to a radius of 115 kpc \citep{Moore06}.  Stellar disks are typically embedded in much larger \HI\ disks.  The size of the GLSB stellar disks is similar to the largest \HI\ disks observed in nearby galaxies, in support of the \citet{Mapelli08} model where the GLSB disks are the faded remnants of a normal \HI\ disk that has experienced a merger induced star-burst.

For comparison:  \citet{Williams13} trace star formation out to 11 disk scale lengths (radius of 22 kpc) in NGC 2403.

@@

In Figure~\ref{disk_comp}, we compare the sizes of the stellar disks of the GLSB galaxies with other observations.  For this plot, we take the maximum detected radius from the literature for the galaxies.  It is possible that deeper observations could increase the values of $R_{max}$.  While UGC 6614 is close in size to the upper end of local \HI\ disks, Malin 1 and 2 are much larger than galaxies observed in the THINGS survey.  This seems to point against the GLSB formation mechanism of  \citet{Mapelli08}, as we do not observe \HI\ disks large enough to convert into GLSB systems without some additional source of angular momentum.

%/astro/store/shared-scratch1/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, Malin 1 from \citet{Moore06}, Malin 2 and UGC 6614 from \citet{Pickering97}.  Here, $R_{max}$ is the maximum radius with an \HI\ detection for the THINGS galaxies and the maximum redius with a stellar flux detection for the GLSB galaxies.  It is interesing to note that the stellar disks of GLSB galaxies are often larger than the \HI\ disks of nearby systems.  \label{disk_comp}}
\end{figure}

@@

%/astro/store/scratch/tmp/yoachim/Research/GLSB/THINGS/guess_projen.py
\begin{figure}
\epsscale{0.5}
\plotone{Plots/things_projenguess.eps}
\epsscale{1}
\caption{Comparison of GLSB \HI\ masses \citep{Pickering97} with \HI\ masses measured in the THINGS survey \citep{Walter2008}.  While the \HI\ masses in the GLSB systems are high, they are consistent with what would be available in a merger with a $\sim$110 km/s galaxy, assuming the merger did not trigger large amounts of star formation and feedback. \label{thingsMHI}}
\end{figure}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_Malin2.eps}{Plots/Malin2mu.eps}
\epsscale{.5}
\plotone{Plots/Malin2profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of Malin 2.  Upper Right:  Image made with the VIRUS-P datacube.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  \label{Malin2_fibers} }
\end{figure}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_UGC06614.eps}{Plots/UGC06614mu.eps}
\epsscale{0.5}
\plotone{Plots/UGC06614profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of UGC 06614.  Upper Right:  Image made with the VIRUS-P datacube.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  We only included fibers beyond the outer ring structure.  \label{ugc_fibers} }
\end{figure}
